Milk intake and stroke mortality in the Japan Collaborative Cohort Study - a Bayesian survival analysis

### 2020-07-08 Paperfolder/Paperfolder.tex
updated introduction done

% Keywords
\keyword{keyword 1; keyword 2; keyword 3 (list three to ten pertinent
keywords specific to the article, yet reasonably common within the
subject discipline.).}

@@

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

@@

Dairy food, especially milk has been recommended to reduce stroke risk
by nearly 7\% for each 200 g increment of daily consumption
\citep{DeGoede2016}. More intuitive interpretation for a decreasing risk
would be possible if we were able to compute the exact probability for
people who had milk intake may had lower hazard of dying from stroke
compared with those who never drank milk at all. For general
public/media reporting, concept of hazard in epidemiological studies
could still sometimes be challenging to be understood or misinterpreted
since it should be formally defined as the probability of the occurrence
of an event at a given time point. Usually, authors of epidemiological
papers would tend to use ``risk'' instead of ``hazard'' or
interchangeably. However, it would still possibly be mixed up with
``risk'' that only contain pure ``probability'' without a point or a
period of time in cross-sectional settings. For better understanding and
interpretation of the data that researchers endeavored to collect,
statistical literature have provided plenty of choices that could help
us better communicate with each other.

@@

Full parametric proportional hazard models under Bayesian framework with
Weibull distribution were fitted using Just Another Gibbs Sampler (JAGS)
program \citep{Plummer2003} version 4.3.0 in R version 4.0.1
\citep{RCT2020}. JAGS program is similar to the OpenBUGS
\citep{Lunn2009} project that uses a Gibbs sampling engine for Markov
Chain Monte Carlo (MCMC) simulation. In the current analysis, we
specified non-informative prior distributions for each of the parameters
in our models (\(\beta_n \sim N(0, 1000)\), and
\(\kappa_{\text{shape}} \sim \Gamma(0.001, 0.001)\)). The
Brooks-Gelman-Rubin diagnostic \citep{Brooks1998} was used to refine the
approximate point of convergence, the point when the ratio of the chains
is stable around 1 and the within and between chain variability start to
reach stability was visually checked. The auto-correlation tool further
identified if convergence has been achieved or if a high degree of
auto-correlation exists in the sample. Then, the number of iterations
discarded as `burn-in' was chosen. All models had a posterior sample
size of 100000 from three separated chains with a ``burn-in'' of 2500
iterations. Posterior means (sd) and 95\% Credible Intervals (CrI) of
the estimated HR were presented for each category of milk intake
frequency taking the ``never'' category as the reference. Posterior
probability that the estimated hazard of dying from stroke for the milk
intake for frequency that higher or equal to ``1-2 times/month'' is
smaller compared with those who chose ``never'' to their milk intake
frequency were calculated as \(P(\text{HR} < 1)\).

The parametric forms of the models fitted in the Bayesian survival
analyses include three models: 1) the crude model, 2) the age-centered
adjusted model, 3) and a model further adjusted for potential
confounders which includes: age (centered, continuous), smoking habit
(never, current, former), alcohol intake (never or past, \(<\) 4
times/week, Daily), body mass index (\(<\) 18.5, \(\geq\) 18.5 and \(<\)
25, \(\geq\) 25 and \(<\) 30, \(\geq\) 30 kg/m\(^2\)), history of
hypertension, diabetes, kidney/liver diseases (yes/no), exercise (more
than 1 hour/week, yes/no), sleep duration (\(<\) 7, \(\geq\) 7 and \(<\)
8, \(\geq\) 8 and \(<\) 9, \(\geq\) 9, hours), coffee intake (never,
\(<\) 3-4 times/week, almost daily), education level (attended school
till age 18, yes/no)

\begin{table}[H]

@@

@@

\caption{\label{tab:tab2}Summary of posterior Hazard Ratios (HR) of mortality from total stroke, different stroke type according to the frequency of milk intake in men (JACC study, 1988-2009).}

@@

@@

  & Never & 1-2 times/Month & 1-2 times/Week & 3-4 times/Week & Almost Daily\\

@@

\hspace{1em}MeanHR (SD) & 1 & 0.89 (0.09) & 0.77 (0.07) & 0.79 (0.07) & 0.90 (0.06)\\

@@

\hspace{1em}MCSE & - & 0.0022 & 0.0019 & 0.0022 & 0.0018\\
\rowcolor{gray!6}  \hspace{1em}Pr(HR < 1) & - & 86.5\% & 99.9\% & 99.7\% & 93.5\%\\

@@

\hspace{1em}MeanHR (SD) & 1 & 0.98 (0.11) & 0.84 (0.08) & 0.86 (0.08) & 0.76 (0.05)\\
\rowcolor{gray!6}  \hspace{1em}95\% CrI & - & (0.79, 1.19) & (0.70, 1.00) & (0.71, 1.02) & (0.66, 0.87)\\
\hspace{1em}MCSE & - & 0.0027 & 0.0022 & 0.0021 & 0.0016\\
\rowcolor{gray!6}  \hspace{1em}Pr(HR < 1) & - & 58.7\% & 97.3\% & 96.1\% & 100.0\%\\

@@

@@

\hspace{1em}MCSE & - & 0.0041 & 0.0036 & 0.0038 & 0.0031\\
\rowcolor{gray!6}  \hspace{1em}Pr(HR < 1) & - & 50.6\% & 93.7\% & 89.6\% & 99.0\%\\

@@

Hemorrhagic stroke & 100 & 42 & 58 & 56 & 176\\

@@

\rowcolor{gray!6}  \hspace{1em}MeanHR (SD) & 1 & 1.03 (0.19) & 0.82 (0.14) & 0.84 (0.15) & 0.97 (0.13)\\
\hspace{1em}95\% CrI & - & (0.70, 1.46) & (0.56, 1.14) & (0.60, 1.17) & (0.75, 1.26)\\
\rowcolor{gray!6}  \hspace{1em}MCSE & - & 0.0066 & 0.0057 & 0.0057 & 0.0062\\
\hspace{1em}Pr(HR < 1) & - & 47.2\% & 88.4\% & 86.3\% & 63.1\%\\

@@

@@

\rowcolor{gray!6}  \hspace{1em}MCSE & - & 0.0091 & 0.0067 & 0.0073 & 0.0076\\
\hspace{1em}Pr(HR < 1) & - & 31.6\% & 79.7\% & 76.6\% & 87.6\%\\

@@

\rowcolor{gray!6}  \hspace{1em}MeanHR (SD) & 1 & 1.14 (0.22) & 0.92 (0.17) & 0.95 (0.18) & 0.95 (0.14)\\
\hspace{1em}95\% CrI & - & (0.75, 1.61) & (0.63, 1.29) & (0.65, 1.37) & (0.71, 1.27)\\
\rowcolor{gray!6}  \hspace{1em}MCSE & - & 0.0093 & 0.0075 & 0.0098 & 0.0082\\
\hspace{1em}Pr(HR < 1) & - & 28.8\% & 72.4\% & 64.4\% & 69.3\%\\

@@

@@

\hspace{1em}MCSE & - & 0.0042 & 0.0037 & 0.0036 & 0.0038\\
\rowcolor{gray!6}  \hspace{1em}Pr(HR < 1) & - & 99.1\% & 99.9\% & 99.7\% & 99.5\%\\

@@

\hspace{1em}MeanHR (SD) & 1 & 0.73 (0.13) & 0.65 (0.10) & 0.70 (0.11) & 0.58 (0.07)\\
\rowcolor{gray!6}  \hspace{1em}95\% CrI & - & (0.49, 1.02) & (0.48, 0.88) & (0.51, 0.94) & (0.46, 0.72)\\
\hspace{1em}MCSE & - & 0.0045 & 0.0035 & 0.0041 & 0.0029\\
\rowcolor{gray!6}  \hspace{1em}Pr(HR < 1) & - & 96.9\% & 99.8\% & 98.9\% & 100.0\%\\

@@

@@

\hspace{1em}MCSE & - & 0.0049 & 0.0047 & 0.0061 & 0.0052\\
\rowcolor{gray!6}  \hspace{1em}Pr(HR < 1) & - & 96.1\% & 99.1\% & 97.5\% & 99.8\%\\

@@

@@

\noindent
\begin{tabular}{@{}ll}
JACC & Japan Collaborative Cohort \\
FFQ & Food Frequency Questionnaire \\
MCMC & Markov Chain Monte Carlo \\
JAGS & Just Another Gibbs Samplers \\
HR & hazard ratio \\
sd & standard deviation \\
CrI & credible interval \\
\end{tabular}}

>>>>

% Keywords
\keyword{milk intake; mortality; stroke; Bayesian survival anlysis;
time-to-event data; JACC study}

@@

\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

@@

Dairy food, especially milk has been recommended to reduce stroke risk
by nearly 7\% for each 200 g increment of daily consumption
\citep{DeGoede2016}. More intuitive interpretation for a decreasing risk
would be possible if we were able to compute the exact probability for
people who had milk intake may had lower hazard of dying from stroke
compared with those who never drank milk at all. For general
public/media reporting, concept of hazard in epidemiological studies
could still sometimes be challenging to be understood or misinterpreted
since hazard is formally defined as the probability of the occurrence of
an event at a given time point \citep{collett2015modelling}. Usually,
authors of epidemiological papers would tend to use ``risk'' instead of
``hazard'' or interchangeably. However, it would still possibly be mixed
up with ``risk'' that only contain pure meaning of ``probability of an
event'' without redefining a point or a period of time in
cross-sectional settings. For better understanding and interpretation of
the findings from data that researchers endeavored to collect,
statistical literature have provided plenty of choices that could help
us better communicate with each other. Another approach of comparing the
time-to-event survival probabilities between different groups would be
to model the time before observing an event rather than the hazard which
always required the assumption of a proportional hazard. Accelerated
failure time models are among these convenient tools that would avoid
frequently unmet assumption of proportional hazard and directly show how
faster/slower that one individual in an exposure group might have an
event compared to others among different exposure groups
\citep{Wei1992}.

Our aim was to overcome these potential pitfalls, avoid
misunderstanding, and provide a more straightforward answer to the main
research question that whether someone answered he/she drank milk at the
baseline of study had lower hazard of dying from stroke compared with
his/her counterparts who said they never consumed milk. If the answer to
the primary objective was yes, then the probabilities that individuals
with different frequencies of milk intake may had lower hazard compared
with those who never drank milk were calculated through a Markov Chain
Monte Carlo (MCMC) simulation process. A Bayesian survival analysis
method was applied on an existing database and through which, we also
provided estimates about whether drinking milk could delay or slow down
the speed towards a stroke event from happening after controlling for
the other potential confounders.

@@

Full parametric proportional hazard models under Bayesian framework with
Weibull distribution were fitted using Just Another Gibbs Sampler (JAGS)
program \citep{Plummer2003} version 4.3.0 in R version 4.0.1
\citep{RCT2020}. JAGS program is similar to the OpenBUGS
\citep{Lunn2009} project that uses a Gibbs sampling engine for MCMC
simulation. In the current analysis, we specified non-informative prior
distributions for each of the parameters in our models
(\(\beta_n \sim N(0, 1000)\), and
\(\kappa_{\text{shape}} \sim \Gamma(0.001, 0.001)\)). The
Brooks-Gelman-Rubin diagnostic \citep{Brooks1998} was used to refine the
approximate point of convergence, the point when the ratio of the chains
is stable around 1 and the within and between chain variability start to
reach stability was visually checked. The auto-correlation tool further
identified if convergence has been achieved or if a high degree of
auto-correlation exists in the sample. Then, the number of iterations
discarded as `burn-in' was chosen. All models had a posterior sample
size of 100000 from four separated chains with a ``burn-in'' of 2500
iterations. Posterior means (sd) and 95\% Credible Intervals (CrI) of
the estimated hazard ratios (HRs) as well as acceleration factors (AFs)
were presented for each category of milk intake frequency taking the
``never'' category as the reference. Posterior probabilities that the
estimated hazard of dying from stroke for the milk intake for frequency
that higher or equal to ``1-2 times/month'' is smaller compared with
those who chose ``never'' to their milk intake frequency were calculated
as \(P(\text{HR} < 1)\).

The parametric forms of the models fitted in the Bayesian survival
analyses included three models: 1) the crude model, 2) the age-centered
adjusted model, 3) and a model further adjusted for potential
confounders which includes: age (centered, continuous), smoking habit
(never, current, former), alcohol intake (never or past, \(<\) 4
times/week, Daily), body mass index (\(<\) 18.5, \(\geq\) 18.5 and \(<\)
25, \(\geq\) 25 and \(<\) 30, \(\geq\) 30 kg/m\(^2\)), history of
hypertension, diabetes, kidney/liver diseases (yes/no), exercise (more
than 1 hour/week, yes/no), sleep duration (\(<\) 7, \(\geq\) 7 and \(<\)
8, \(\geq\) 8 and \(<\) 9, \(\geq\) 9, hours), coffee intake (never,
\(<\) 3-4 times/week, almost daily), education level (attended school
till age 18, yes/no)

\begin{table}[h]

@@

\end{tabular}}

@@

\caption{\label{tab:tab2}Summary of posterior Acceleration Factors (AF) and Hazard Ratios (HR) of mortality from total stroke, stroke types according to the frequency of milk intake in men (JACC study, 1988-2009).}

@@

\fontsize{6}{8}\selectfont
\resizebox*{\textwidth}{\dimexpr\textheight-\lineskip\relax}{%

@@

\textbf{ } & \textbf{Never} & \textbf{1-2 times/Month} & \textbf{1-2 times/Week} & \textbf{3-4 times/Week} & \textbf{Almost Daily}\\

@@

\hspace{1em}Mean AF (SD) & 1 & 0.93 (0.07) & 0.83 (0.05) & 0.85 (0.05) & 0.93 (0.04)\\
\rowcolor{gray!6}  \hspace{1em}95\% CrI & - & (0.81, 1.06) & (0.73, 0.94) & (0.74, 0.96) & (0.85, 1.02)\\
\hspace{1em}Mean HR (SD) & 1 & 0.89 (0.09) & 0.77 (0.07) & 0.79 (0.07) & 0.90 (0.06)\\

@@

\hspace{1em}Pr(HR < 1) & - & 86.50\% & 99.90\% & 99.70\% & 93.50\%\\

@@

\rowcolor{gray!6}  \hspace{1em}Mean AF (SD) & 1 & 0.99 (0.06) & 0.90 (0.05) & 0.91 (0.05) & 0.85 (0.04)\\
\hspace{1em}95\% CrI & - & (0.87, 1.11) & (0.81, 1.00) & (0.82, 1.01) & (0.78, 0.92)\\
\rowcolor{gray!6}  \hspace{1em}MeanHR (SD) & 1 & 0.98 (0.11) & 0.84 (0.08) & 0.86 (0.08) & 0.76 (0.05)\\
\hspace{1em}95\% CrI & - & (0.79, 1.19) & (0.70, 1.00) & (0.71, 1.02) & (0.66, 0.87)\\
\rowcolor{gray!6}  \hspace{1em}Pr(HR < 1) & - & 58.70\% & 97.30\% & 96.10\% & 100.00\%\\

@@

\hspace{1em}Mean AF (SD) & 1 & 1.00 (0.07) & 0.92 (0.06) & 0.94 (0.06) & 0.88 (0.05)\\
\rowcolor{gray!6}  \hspace{1em}95\% CrI & - & (0.88, 1.14) & (0.82, 1.03) & (0.84, 1.05) & (0.81, 0.96)\\

@@

\hspace{1em}Pr(HR < 1) & - & 50.60\% & 93.70\% & 89.60\% & 99.00\%\\

@@

\rowcolor{gray!6}  Hemorrhagic stroke & 100 & 42 & 58 & 56 & 176\\

@@

\hspace{1em}Mean AF (SD) & 1 & 1.03 (0.17) & 0.85 (0.12) & 0.87 (0.13) & 0.98 (0.11)\\
\rowcolor{gray!6}  \hspace{1em}95\% CrI & - & (0.74, 1.38) & (0.63, 1.12) & (0.65, 1.14) & (0.78, 1.22)\\
\hspace{1em}MeanHR (SD) & 1 & 1.03 (0.19) & 0.82 (0.14) & 0.84 (0.15) & 0.97 (0.13)\\
\rowcolor{gray!6}  \hspace{1em}95\% CrI & - & (0.70, 1.46) & (0.56, 1.14) & (0.60, 1.17) & (0.75, 1.26)\\
\hspace{1em}Pr(HR < 1) & - & 47.20\% & 88.40\% & 86.30\% & 63.10\%\\

@@

\rowcolor{gray!6}  \hspace{1em}Mean AF (SD) & 1 & 1.08 (0.17) & 0.91 (0.13) & 0.92 (0.13) & 0.90 (0.10)\\
\hspace{1em}95\% CrI & - & (0.80, 1.45) & (0.70, 1.20) & (0.71, 1.19) & (0.74, 1.11)\\

@@

\rowcolor{gray!6}  \hspace{1em}Pr(HR < 1) & - & 31.60\% & 79.70\% & 76.60\% & 87.60\%\\

@@

\hspace{1em}Mean AF (SD) & 1 & 1.11 (0.18) & 0.93 (0.15) & 0.96 (0.16) & 0.96 (0.13)\\
\rowcolor{gray!6}  \hspace{1em}95\% CrI & - & (0.79, 1.58) & (0.70, 1.25) & (0.71, 1.34) & (0.76, 1.25)\\
\hspace{1em}MeanHR (SD) & 1 & 1.14 (0.22) & 0.92 (0.17) & 0.95 (0.18) & 0.95 (0.14)\\
\rowcolor{gray!6}  \hspace{1em}95\% CrI & - & (0.75, 1.61) & (0.63, 1.29) & (0.65, 1.37) & (0.71, 1.27)\\
\hspace{1em}Pr(HR < 1) & - & 28.80\% & 72.40\% & 64.40\% & 69.30\%\\

@@

\hspace{1em}Mean AF (SD) & 1 & 0.76 (0.09) & 0.71 (0.07) & 0.74 (0.08) & 0.79 (0.06)\\
\rowcolor{gray!6}  \hspace{1em}95\% CrI & - & (0.59, 0.94) & (0.58, 0.86) & (0.61, 0.89) & (0.68, 0.93)\\

@@

\hspace{1em}Pr(HR < 1) & - & 99.10\% & 99.90\% & 99.70\% & 99.50\%\\

@@

\rowcolor{gray!6}  \hspace{1em}Mean AF (SD) & 1 & 0.83 (0.08) & 0.79 (0.07) & 0.82 (0.07) & 0.74 (0.05)\\
\hspace{1em}95\% CrI & - & (0.68, 1.01) & (0.67, 0.93) & (0.69, 0.96) & (0.66, 0.84)\\
\rowcolor{gray!6}  \hspace{1em}MeanHR (SD) & 1 & 0.73 (0.13) & 0.65 (0.10) & 0.70 (0.11) & 0.58 (0.07)\\
\hspace{1em}95\% CrI & - & (0.49, 1.02) & (0.48, 0.88) & (0.51, 0.94) & (0.46, 0.72)\\
\rowcolor{gray!6}  \hspace{1em}Pr(HR < 1) & - & 96.90\% & 99.80\% & 98.90\% & 100.00\%\\

@@

\hspace{1em}Mean AF (SD) & 1 & 0.84 (0.09) & 0.80 (0.08) & 0.83 (0.08) & 0.75 (0.06)\\
\rowcolor{gray!6}  \hspace{1em}95\% CrI & - & (0.67, 1.02) & (0.67, 0.95) & (0.69, 0.99) & (0.66, 0.85)\\

@@

\hspace{1em}Pr(HR < 1) & - & 96.10\% & 99.10\% & 97.50\% & 99.80\%\\

@@

\fontsize{7}{9}\selectfont

@@

\noindent
\begin{tabular}{@{}ll}
JACC & Japan Collaborative Cohort \\
FFQ & Food Frequency Questionnaire \\
MCMC & Markov Chain Monte Carlo \\
JAGS & Just Another Gibbs Samplers \\
AFT & accelerated failure time \\
HR & hazard ratio \\
AF & acceleration factor \\
sd & standard deviation \\
CrI & credible interval \\
\end{tabular}}
